Reproducible research in archaeology using R & rrtools

### 2018-01-26 analysis/paper/paper.Rmd
complete text mining of titles

library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

@@

```{r}
library(tidyverse)

wc_archaeology_2017 <- 
  readr::read_tsv(paste0(here::here(), "/analysis/data/raw_data/savedrecs.txt"), 
            quote = "")
```

@@

library(ggrepel)
ggplot(most_popular_journals_over_time,
       aes(PY, 
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(most_popular_journals_over_time, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 15,
    segment.color = NA
  ) +
  xlim(2007, 2019) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of articles per year in Archaeology journals",
          subtitle = "Data from apps.webofknowledge.com,\nusing WC=Archaeology in SCI-EXPANDED & SSCI for 2007-2017")
```

@@

We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`

@@

cran_cites_files <- 
    paste0(here::here(), glue("/analysis/data/raw_data/savedrecs ({3:11}).txt"))
```

```{r fig.height=10}
cran_cites_all_areas_all_years <- 
  map(cran_cites_files,
      ~readr::read_tsv(.x, 
                       quote = "",
                       col_types = cols(.default = col_character()))) %>% 
  bind_rows()

@@

# tally by year
cran_cites_all_areas_top_journals_by_year <- 
  cran_cites_all_areas_all_years %>% 
  filter(SO %in% top_journals_for_cran_cites$SO[1:10]) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  ungroup() %>% 
  mutate(PY = as.numeric(PY))

ggplot(cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(cran_cites_all_areas_top_journals_by_year, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 1,
    segment.color = NA
  ) +
  xlim(2007, 2018) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of articles per year citing R Core in top 10 WoS journals",
          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
```

@@

ggplot(cran_cites_in_archy_journals_per_year,
       aes(PY,
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(cran_cites_in_archy_journals_per_year, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 1,
    segment.color = NA
  ) +
  xlim(2007, 2018) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of Archaeology articles per year citing R Core",
          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
```

@@

```{r eval = FALSE}
library(tidytext)

cran_cites_in_archy_journals %>% 
  select(TI) %>% 
  unnest_tokens(word, TI) %>% 
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>% 
  head(20)

library(textrank)

# keywords for each title
keyw <- map(cran_cites_in_archy_journals$TI, 
            ~textrank_keywords(unlist(str_split(paste0(.x, collapse = " "), " "))[!unlist(str_split(paste0(.x, collapse = " "), " ")) %in% stop_words$word]))

map(keyw, ~enframe(.x$pagerank$vector) %>% 
  arrange(desc(value)))

# tf-idf to show words most important to each journal
# not very interesting
cran_cites_in_archy_journals %>% 
  filter(SO %in% top_journals_by_articles$SO) %>% 
  select(SO, TI) %>% 
  unnest_tokens(word, TI) %>% 
  anti_join(stop_words) %>%   
  count(SO, word, sort = TRUE) %>% 
  bind_tf_idf(word, SO, n) %>% 
  arrange(desc(tf_idf)) %>% 
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(SO) %>% 
  top_n(5) %>% 
  ungroup %>%
  ggplot(aes(word, tf_idf, fill = SO)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~SO, ncol = 2, scales = "free") +
  coord_flip()
```

>>>>

library(caa2018) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory

library(ggrepel)
library(tidyverse)
```

@@

```{r}
wc_archaeology_2017 <- 
  readr::read_tsv(paste0(here::here(), "/analysis/data/raw_data/savedrecs.txt"), 
            quote = "")
```

@@

ggplot(most_popular_journals_over_time,
       aes(PY, 
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(most_popular_journals_over_time, 
                  PY == max(PY)),
    aes(label = SO),
    size = 3,
    nudge_x = 15,
    segment.color = NA
  ) +
  xlim(2007, 2019) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of articles per year in Archaeology journals",
          subtitle = "Data from apps.webofknowledge.com,\nusing WC=Archaeology in SCI-EXPANDED & SSCI for 2007-2017")
```

@@

We did a 'Cited Reference Search' in WoS for title "R: A Language and Environment for Statistical Computing" and author = "R Core Team", which is the output from  `citation()`. This function first appeared in 2004 https://github.com/wch/r-source/commits/9a5c086855d2e8fb07f670e3315871b2de31a4fd/src/library/base/inst/CITATION

Note that before 2012 the authors were "R Development Core Team", and then it was simplified to "R Core Team" https://github.com/wch/r-source/commit/9a5c086855d2e8fb07f670e3315871b2de31a4fd#diff-80bb8a8b133a6204e9642d76791927e2

This is our exact search string for WoS:

You searched for: CITED AUTHOR: ("R DEV COR TEAM" OR "R CORE" OR "R CORE TEAM") AND CITED WORK: ("R: A Language and Environment for Statistical Computing") AND LANGUAGE: (English) AND DOCUMENT TYPES: (Article)
Timespan: All years. Indexes: SCI-EXPANDED, SSCI, A&HCI, ESCI.

@@

cran_cites_files <- 
    paste0(here::here(), glue("/analysis/data/raw_data/cited_cran/savedrecs ({1:11}).txt"))
```

```{r fig.height=10}
cran_cites_all_areas_all_years <- 
  map(cran_cites_files,
      ~readr::read_tsv(.x, 
                       quote = "",
                       col_types = cols(.default = col_character()))) %>% 
  bind_rows() %>% 
  mutate(PY = as.numeric(PY)) %>% 
  filter(PY <= 2017)

@@

# tally by year
cran_cites_all_areas_top_journals_by_year <- 
  cran_cites_all_areas_all_years %>% 
  filter(SO %in% top_journals_for_cran_cites$SO[1:10]) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  ungroup()

ggplot(cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(cran_cites_all_areas_top_journals_by_year, 
                  PY == max(PY)),
    aes(label = str_wrap(SO, 15)),
    size = 3,
    direction = "both",
    nudge_x = 1,
    segment.color = NA
  ) +
  xlim(2007, 2018) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of articles per year citing R Core in top 10 WoS journals",
          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
```

@@

ggplot(cran_cites_in_archy_journals_per_year,
       aes(PY,
           n,
           colour = SO)) +
  geom_line() +
  geom_text_repel(
    data = subset(cran_cites_in_archy_journals_per_year, 
                  PY == max(PY)),
    aes(label = str_wrap(SO, 10)),
    size = 3,
    nudge_x = 1,
    segment.color = NA
  ) +
  xlim(2007, 2019) +
  theme_minimal() +
  theme(legend.position="none") +
  ggtitle("Number of Archaeology articles per year citing R Core",
          subtitle = "Data from apps.webofknowledge.com,\nusing in SCI-EXPANDED & SSCI for 2007-2017")
```

@@

```{r}
library(tidytext)
# https://www.tidytextmining.com/twitter.html

# These are all archy articles
# wc_archaeology_2007_to_2017

# These are ones citing R
# cran_cites_in_archy_journals

# subtract ones citing R from ones that do not
library(Hmisc)
wc_archaeology_2007_to_2017_no_R_cites <- 
  wc_archaeology_2007_to_2017 %>% 
  filter(TI %nin% cran_cites_in_archy_journals$TI)

# combine
archy_journal_articles <- 
  bind_rows(wc_archaeology_2007_to_2017_no_R_cites %>% 
            mutate(ID = "no_R_cites",
                   VL = as.character(VL)),
           cran_cites_in_archy_journals %>% 
            mutate(ID = "has_R_cites"))

library(tidytext)
library(stringr)

tidy_cites <- 
archy_journal_articles %>% 
  unnest_tokens(word, TI) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "[a-z]"))

frequency <- 
  tidy_cites %>% 
  group_by(ID) %>% 
  count(word, sort = TRUE) %>% 
  left_join(tidy_cites %>% 
              group_by(ID) %>% 
              summarise(total = n())) %>%
  mutate(freq = n/total)

frequency <- frequency %>% 
  select(ID, word, freq) %>% 
  spread(ID, freq) %>%
  arrange(no_R_cites, has_R_cites)

library(scales)

ggplot(frequency, 
       aes(no_R_cites, 
           has_R_cites)) +
  geom_jitter(alpha = 0.1, 
              size = 2.5, 
              width = 0.25, 
              height = 0.25) +
  geom_text(aes(label = word), 
            check_overlap = TRUE, 
            vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

word_ratios <- tidy_cites %>%
  count(word, ID) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(ID, n, fill = 0) %>%
  mutate_if(is.numeric, funs((. + 1) / sum(. + 1))) %>%
  mutate(logratio = log(has_R_cites / no_R_cites)) %>%
  arrange(desc(logratio))

word_ratios %>%
  group_by(logratio < 0) %>%
  top_n(10, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (has_R_cites/no_R_cites)") +
  scale_fill_discrete(name = "", labels = c("has_R_cites", "no_R_cites"))

```
